### 2022-09-20 main/chapter_2ljets.tex
add ssfm cite; use auto bib for foundations and algorithms
more awful attempts at writing sm measurements; fix - in pages

The magnitude of $\ptmiss$ is $\met$, pronounced `met'.
Note that despite its $E$, $\met$ is not an energy; transverse energy does not exist.
One may, of course, define a directed `transverse energy', but it will
not be conserved (in the standard model) unless it is equivalent to a conserved
momentum.
Indeed, momentum-balanced events do have energy imbalances;
for example, in a $Z\mathrm{+jet}$ event $pp \to Zg$, with the $Z$ and
$g$ emitted back-to-back in the transverse plane with equal momenta
$p = |\pm \vec p|$, the two particles have different energies since they
have different masses: $E_Z^2 = m_Z^2 + p^2$, $E_g^2 = 0 + p^2$.
% hobbyhorse

@@

\begin{figure}[tp]
\centering
\includegraphics[width=0.9\textwidth]{figures/2ljets_compressed_3l_ins1866951_fig_16a.png}
\caption[
Contours showing off-shell C1N2 exclusion from the \atlas\ Run-2 $3\ell$ search
]{%
Contours showing off-shell C1N2 exclusion in the three lepton channel,
reproduced from the \atlas\ Run-2 $3\ell$
search~\cite{atlas_rjr_3l_SUSY_2019_09, hepdata.95751}.
Off-shell models are excluded in that large lobe squeezed between the
$m(\neutralino_2) = m(\neutralino_1)$ and
$m(\neutralino_2) = m(\neutralino_1) - m_Z$ diagonal lines.
}
\label{fig:ljets_offshell_3l_exclusion}
\end{figure}

@@

% pro
Naturally, \atlas\ has an unblinding procedure that requires committee
review and approval in accordance with internal documentation.
In the \atlas\ SUSY world, searches are nominally performed blind to signal
regions.
For these searches, their control and validation regions are used in the
design and modelling of signal regions, so they must be inspected.
This is not a problem since control and validation regions should also
have negligibly small yields (contaminations) from interesting signal models,
With negligible signal contaminations, control and validation region data can
be consumed to constrain the background modelling, in composition with all
other data on which the Standard Model stands, and without being influenced by
possible data from the target signal models that are to be tested in signal
regions.

>>>>

The magnitude of $\ptmiss$ is $\met$, pronounced `met'.
Note that despite its $E$, $\met$ is not an energy; transverse energy does not exist.
One may, of course, define a directed `transverse energy', but it will
not be conserved (in the standard model) unless it is equivalent to a conserved
momentum.
Indeed, momentum-balanced events do have energy imbalances;
for example, in a $Z\mathrm{+jet}$ event $pp \to Zg$, with the $Z$ and
$g$ emitted back-to-back in the transverse plane with equal momenta
$p = |\pm \vec p|$, the two particles have different energies since they
have different masses: $E_Z^2 = m(Z)^2 + p^2$, $E_g^2 = 0 + p^2$.
% hobbyhorse

@@

\begin{figure}[tp]
\centering
\includegraphics[width=0.9\textwidth]{figures/2ljets_compressed_3l_ins1866951_fig_16a.png}
\caption[
Contours showing off-shell C1N2 exclusion from the \atlas\ Run-2 $3\ell$ search
]{%
Contours showing off-shell C1N2 exclusion in the three lepton channel,
reproduced from the \atlas\ Run-2 $3\ell$
search~\cite{atlas_rjr_3l_SUSY_2019_09, hepdata.95751}.
Off-shell models are excluded in that large lobe squeezed between the
$m(\neutralino_2) = m(\neutralino_1)$ and
$m(\neutralino_2) = m(\neutralino_1) - m(Z)$ diagonal lines.
}
\label{fig:ljets_offshell_3l_exclusion}
\end{figure}

@@

% pro
Naturally, \atlas\ has an unblinding procedure that requires committee
review and approval in accordance with internal
documentation~\cite{usa1944simple}.
In the \atlas\ SUSY world, searches are nominally performed blind to signal
regions.
For these searches, their control and validation regions are used in the
design and modelling of signal regions, so they must be inspected.
This is not a problem since control and validation regions should also
have negligibly small yields (contaminations) from interesting signal models,
With negligible signal contaminations, control and validation region data can
be consumed to constrain the background modelling, in composition with all
other data on which the Standard Model stands, and without being influenced by
possible data from the target signal models that are to be tested in signal
regions.

### 2022-09-20 main/chapter_context.tex
more awful attempts at writing sm measurements; fix - in pages

The \atlas\ detector was not delivered in a preordained form.
It was designed with intent to provide maximally useful and novel data within
real-world cost budgets.
Original designs hoped to construct an \atlas\ detector that could
observe the Higgs boson (which was predicted but yet unseen),
precisely measure known massive objects ($W$, $Z$, and $t$),
and possibly discover new particles such as those hypothesized by
supersymmetric models~\cite{atlas1999design1}.
We now review these three goals with the subsequent news from \atlas'
operations.

\paragraph{Higgs boson discovery.}
Prior to \atlas' construction, the Higgs boson mass (and existence) was
poorly known, but loosely constrained by results like
$m_H \gtrsim 90\,\eV[G]$ from searches at the
Large Electron–Positron Collider (LEP),
$m_H = 76^{+85}_{-47}\,\eV[G]$ from other data on the electroweak sector,
and $m_H \lesssim 1\,\eV[T]$ from theoretical unitarity arguments~\cite{
atlas1999design2,
ghinculov1998perturb,
lep1999ewk
}.
The Higgs boson couples to an extraordinary variety of fundamental particles.
For its discovery alone, \atlas' designers achieved sensitivity across the
entire $100$ to $1000\,\eV[G]$ window by targetting decays including
$H\to b\bar b$,
$H\to \gamma\gamma$,
$H\to ZZ^* \to 4\ell$, and
$H\to ZZ \to \ell\ell\nu\nu$;
these channels demand precise measurements of leptons, photons, and
$b$-jets, as well as sensitivity to the missing transverse momentum of
invisible neutrinos for $ZZ \to \ell\ell\nu\nu$ decays at the very highest
masses.

Both \atlas\ and \cms\ have now found a Higgs boson candidate at
$m_H = 125\,\eV[G]$~\cite{
atlas2012higgs,
atlas2012combined,
cms2012higgs
}
and confirmed all of its tested Standard Model properties~\cite{
combined2016higgs,
atlas2022ten,
cms2022ten
}.

\paragraph{Measurement of known massive objects.}
Other than the Higgs boson, the heavy Standard Model resonances are the
$W$ and $Z$ bosons and the top quark, all of which are produced at
high rates by LHC proton-proton collisions.

The goal of measuring these heavy objects puts more emphasis on determining
the total energy of hadronic jets than their precise hadronic compositions.

\TODO{Although \atlas\ has an active B-physics programme, the stronger particle
identification of the \lhcb\ detector makes it more competitive for hadronic
physics.?}
\TODO{Heavy ions and ALICE?}

@@

@@

\paragraph{Supersymmetry discovery.}
This is not a failure. It is a success.
Our goal is not to discover something that is not there, but to learn more
about nature.
The nonappearance of supersymmetric particles at the LHC is itself a discovery.

\paragraph{Design.}
Atlas is a Titan.
On the \cern\ Large Hadron Collider (LHC), \atlas\ is the largest detector
experiment, and alongside its majors neighbours
\cms~\cite{cms2008experiment},
\lhcb~\cite{lhcb2008experiment},
and ALICE~\cite{alice2008experiment}
it is a multi-kilotonne beast,
and a wonder of scientific engineering~\cite{
atlas2008experiment,
atlas1999design1,
atlas1999design2
}.

>>>>

The \atlas\ detector was not delivered in a preordained form.
It was designed with intent to provide maximally useful and novel data within
real-world cost budgets.
Within general purpose utilities, \atlas' designs aimed to construct a
detector that could
\begin{itemize}
\item find the Higgs boson (which was predicted but yet unseen),
\item precisely measure known massive objects ($W$, $Z$, and $t$),
\item discover new (supersymmetric) particles~\cite{atlas1999design1}.
\end{itemize}
We now review these three goals with the subsequent news from \atlas'
operations.

\paragraph{Find the Higgs boson.}
Prior to \atlas' construction, the Higgs boson mass (and existence) was
poorly known, but loosely constrained by results like
$m(H) \gtrsim 90\,\eV[G]$ from searches at the
Large Electron–Positron Collider (LEP),
$m(H) = 76^{+85}_{-47}\,\eV[G]$ from other data on the electroweak sector,
and $m(H) \lesssim 1\,\eV[T]$ from theoretical unitarity arguments~\cite{
atlas1999design2,
ghinculov1998perturb,
lep1999ewk
}.

The Higgs boson couples to an extraordinary variety of fundamental particles.
For its discovery alone, \atlas' designers achieved sensitivity across the
entire $100$ to $1000\,\eV[G]$ window by targetting decays including
$H\to b\bar b$,
$H\to \gamma\gamma$,
$H\to ZZ^* \to 4\ell$, and
$H\to ZZ \to \ell\ell\nu\nu$;
these channels demand precise measurements of leptons, photons, and
$b$-jets, as well as sensitivity to the missing transverse momentum of
invisible neutrinos for $ZZ \to \ell\ell\nu\nu$ decays at the very highest
masses.

Both \atlas\ and \cms\ have now found a Higgs boson candidate at
$m(H) = 125\,\eV[G]$~\cite{
atlas2012higgs,
atlas2012combined,
cms2012higgs
}
and confirmed all of its tested Standard Model properties~\cite{
combined2016higgs,
atlas2022ten,
cms2022ten
}
such as its production mechanisms, decay rates,
spin, and parity~\cite{HIGG-2013-01, HIGG-2013-17, HIGG-2014-06}.

\paragraph{Precisely measure known massive objects.}
Other than the Higgs boson, the heavy Standard Model resonances are the
$W$ and $Z$ bosons and the top quark.
All of these are produced at high rates in LHC proton-proton collisions,
and present opportunities for \atlas\ to measure their properties
(such as masses and production cross-sections) with unprecedented precision.

Sensitivity to hadronic momenta is crucial,
both directly to measure heavy hadronic decays
(like $t \rightarrow bq\bar{q}$),
and indirectly for the missing transverse momenta from semi-invisible
leptonic decays (like $W \to \ell\nu$).
Since individual hadron identities are less relevant to these measurements,%
\footnote{%
Rare decays like
$W^\pm\rightarrow \pi^\pm \gamma$ or $W^\pm\rightarrow \pi^\pm \pi^+ \pi^-$
could give clean signals, but are not competitive due to their
tiny production rates and difficult identification~\cite{
cdf1996search,
mangano2014wpiy,
cms2021wpiy
}.%
}
it makes sense for \atlas' design to forfeit hadron identification systems
in favour of precise calorimetry with total angular coverage.

@@

\TODO{w mass conflict~\cite{cdf2022high}}

\paragraph{Discover new (supersymmetric) particles.}
This is not a failure. It is a success.
Our goal is not to discover something that is not there, but to learn more
about nature.
The nonappearance of supersymmetric particles at the LHC is itself a discovery.

\TODO{
Decay to SM particles,
emphasis on missing transverse momentum.
...
Many SUSY searches, many other exotics searches, no convincing evidence for
new effects.
}

@@

\subsection{Design}
Atlas is a Titan.
On the \cern\ Large Hadron Collider (LHC), \atlas\ is the largest detector
experiment, and alongside its majors neighbours
\cms~\cite{cms2008experiment},
\lhcb~\cite{lhcb2008experiment},
and ALICE~\cite{alice2008experiment}
it is a multi-kilotonne beast,
and a wonder of scientific engineering~\cite{
atlas2008experiment,
atlas1999design1,
atlas1999design2
}.

### 2022-09-20 main/draft.tex
more awful attempts at writing sm measurements; fix - in pages

>>>>

% awful awful tangled up day

\TODO{B-hadrons and LHCb?}
% more cost-effective to improve calorimetry at the expense of PID

% measurements are supported by complete and precise calorimetry
% PID is expensive and less useful in comparison

To precise measure of these massive objects
justifies a focus on complete and precise calorimetry
at the expense of exact particle identification.

\TODO{
Argument:
want to measure W, t,
so measure jets and precise objects
so weaker B and heavy ion programme.
...
Refer to competitive measurements of W and t masses
and xsec of various rare processes (diboson, triboson)
}

Towards maximizing the quality of these precision measurements,

sensitivity to
exact hadron identities is comparatively less cost-effective.%

To maximize the quality of measurements like $m(t)$ and $m(W)$, it therefore
makes sense for \atlas' design to prefer complete and accurate calorimetry over
hadron identification.

Such hadronic momentum measurements contribute , both of which are
leveraged for \atlas' measurements of $m(t)$ and $m(W)$~\cite{
atlas2022symmarytop,
atlas2019topmass
}.

Although \atlas\ has an active programmes in other Standard Model research such
as B-physics and heavy-ion collisions, these are rarely competitive with the specialized experiments
\lhcb\ and ALICE respectively.
